---
title: "lab4_part1 - CI and t-tests"
author: "Satie Airame"
date: "10/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## 1. Confidence intervals

```{r}
otter_length <- c(38, 41, 50, 27, 32, 41, 48, 60, 43)

t.test(otter_length)
sd(otter_length)

```

Meaningful mean and CI but t-test outcome is not meaningful.
Tests null hypothesis the default that mean otter length is 0 inches. Extremely unlikely, does not make sense.

## T-test (1-sample t-test)

See a claim: mean otter length is 50 inches.
mu = null hypthesis that I am testing

```{r}
otter_test <- t.test(x = otter_length, mu = 50)
otter_test # To have show up in Markdown document
```
p-value of 0.0444 
If population mean really is 50 inches, then there is a 4.4% chance that I could have randomly taken a sample from that population that has a mean that is at least a different from 50 as my sample mean is in either direction.
Take away: Is it possible that my sample was drawn from a sample that has a mean of 50. Yes, possible but only a 4.44% chance that I could have drawn a random sample and found a mean that is at least as different as mine is from that null hypothesis.

## Two-sample t-test
 
```{r}
desert_bighorns <- c(32, 44, 18, 26, 50, 33, 42, 40)
sierra_bighorns <- c(28, 31, 40, 42, 26, 29, 31)

t.test(x = desert_bighorns, y = sierra_bighorns)
```

Always run histogram or qq-plot to see if data are normally distributed.
Welch's Two Sample t-test: does not assume equal variances.

Mean of x and y sample in order that we put vectors into t-function.
CI: What does this mean, how can it be negative. Excpected CI for difference in mean sizes in horn length, y-x can be negative
Difference in mean horn lengths to be within the confidence interval for 95% of samples taken. CI on differences in means not usually reported.
Null hypothesis: difference in means is equal to 0. 
df: sample size (with other things mixed in)
t-statistic: how far away from 0 you are on t-distribution, standardized measure of how far away the difference in your sample means is away from the null hypothesis that difference between the samples means equal to 0.
p-value: probability that if these samples are drawn from populations with the same mean horn length, null hypothesis, then there is almost a 90% chance that I could have taken two random samples from those populations and found samples at least as different as samples I have here. Really good chance that samples are from populations with same mean. Not enough evidence to reject null hypothesis that drawn from .

Remember: p-value is probability if  null hypothesis is true
For two sample t-test, it is the probability if samples were drawn from populations with the same mean value that we could get samples that are at least as different as ours are.
High probability: likely to be drawn from samples from populations with the same mean.
Low probability: less likely to randomly draw samples from populations with the same mean

Focus on actual difference in mean, % difference in mean, what does this mean in the real world.

In Git
M = modified
A = added





